```{r setup and tests, include=FALSE}

## README FIRST ##
#Read and run this chunk of code line by line in the R Console (do not press knit) - there are some questions below which you will have to answer and some logic tests to complete. Once you are happy that the conditions have been satisfied, hit 'knit' above. 

# Load required packages

library(leaflet)
library(dplyr)
library(colortools)
library(corrplot)
library(lubridate)
library(sf)
library(tmap)
library(stringr)
library(tidyr)


# Load your data [change the files paths to your data locations]
# img - image data (including blank images)
# dat - detection data (img, but excluding blank images)
# eff - deployment data (effort)
# sta - station covariates
# ind.dat - independent detection data
# tuyeta - shapefile outline of the study area
# spatdata - independent detections joined spatially with lat longs

project<-"Tuyeta"
independent <- 30 # what threshold did you use (in minutes) for independent detections?

img <- read.csv("../1. Master data/Tuyeta_CAM_Detection_Data_v8.csv", header=T)
img$species_rank<-as.factor(img$species_rank)
img$common_name<-as.factor(img$common_name)
summary(img$species_rank)
summary(img$common_name)
# drop detections that aren't ID-ed to at least the species level:
dat<-droplevels(img[img$species_rank %in% c("Species","Subspecies"),])


eff <- read.csv("../1. Master data/Tuyeta_CAM_Deployment_Data_v8.csv", header=T)

sta <- read.csv("../1. Master data/Tuyeta_Station_Covariates.csv", header=T)
str(sta)

ind.dat <- read.csv("../1. Master data/Tuyeta_CAM_Independent_Detections_v8.csv", header=T)
str(ind.dat)
unique(dat$common_name)
ind.dat<-droplevels(ind.dat[ind.dat$common_name %in% unique(dat$common_name),])


tuyeta <- st_read("C:/Users/laura/Documents/Wildco/5. Spatial/Tuyeta/Tuyeta_boundary/Tuyeta.shp")
tuyeta <- st_transform(tuyeta, crs = 3580)

spatdata=st_read("../1. Master data/Tuyeta_Station_Covariates.csv",
        options=c("X_POSSIBLE_NAMES=longitude","Y_POSSIBLE_NAMES=latitude"))

spatdata<- merge(spatdata, ind.dat, by = "location")

spatdata$common_name<-as.factor(spatdata$common_name)
spatdata$location<-as.factor(spatdata$location)


# Timezone [Use UTC if your cameras do not correct for daylight saving time, if they do use the timezone where the data was collected]
tz <- "UTC"

# Set a single categorical variable of interest from station covariates for summary graphs. If you do not have and appropriate category use "Project.ID".
category <- "project"

# Define a colour from the R options to base the colourscheme
colour <- "lightseagreen"



## DATA TESTS ##


# This code will not work unless your data passes the following checks
 
# 2) All dates must be in YYYY-MM-DD in 'eff' and YYYY-MM-DD HH:MM:SS in 'dat' 
# If the following return NA, change your formatting
strptime(eff$deployment_begin_date[1], "%Y-%m-%d", tz="UTC")
strptime(dat$date_detected[1], "%Y-%m-%d %H:%M:%S", tz="UTC")

# 3) the dates in 'eff$deployment_end_date' must be the when the camera fails, not when you check the camera. If the camera fails (due to damage or full sd card), use the last day it functions here.  

# 4) Ensure your species names are consistent - check in the list below
table(dat$common_name)

# 5) Ensure Number.of.Animals doesn't have any non-numeric data in! The following should return TRUE
is.numeric(dat$count)

# 6) ensure all deployment dates are before retrieval dates for each deployment
# Logic = are the stations active for 0 or more days -> all should read TRUE
table((strptime(eff$deployment_end_date, "%Y-%m-%d", tz="UTC")-strptime(eff$deployment_begin_date, "%Y-%m-%d", tz="UTC"))>=0)

# 7) Do you have lat/long data for all of your sites you have effort data for? If yes, the value should be 0
length(setdiff(eff$location, sta$location))
# If length > 0, then you have some data missing!

# If all of the above is satisfied -> press 'Knit' above ^

```

```{r non-adjustable options, echo=F, include=F}


# Prepare dates
eff$deployment_begin_date <- as.POSIXct(strptime(eff$deployment_begin_date, "%Y-%m-%d", tz ))
eff$deployment_end_date <- as.POSIXct(strptime(eff$deployment_end_date, "%Y-%m-%d", tz ))

eff$Days <- as.numeric(round(difftime(eff$deployment_end_date, eff$deployment_begin_date, units="days"),1))

dat$date_detected <- as.POSIXct(strptime(dat$date_detected, "%Y-%m-%d %H:%M:%S", tz ))

# Count the number of camera stations
n.stat <- length(unique(eff$location))
n.stat

# Generate colours to display the category levels - R needs them as a factor
sta$project <- as.factor(sta$project)
col.cat <- wheel(colour, num = length(levels(sta[,"project"])))
sta$Cols <- col.cat[sta[,"project"]]

# How big should the figures be
eff.height <- 8
if(length(unique(eff$location))>80)
   {
     eff.height <- length(unique(eff$location))/10
   }

sp.height <- 7
if(length(unique(dat$Species))>20)
   {
     sp.height <- 7+(length(unique(dat$Species))/8)
   }


```

# `r project` Project

### Camera locations

To date there have been camera deployments at `r n.stat` unique locations.

```{r map, echo=F}

m <- leaflet() %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data
  addProviderTiles(providers$Esri.WorldTopoMap, group="Base") %>%     
  addCircleMarkers(lng=sta$longitude, lat=sta$latitude,
                   color=sta$Cols,
                   popup=paste(sta$location, sta[,"project"])) %>%
 addLegend("bottomleft", colors = col.cat,  labels = levels(sta[,"project"]),
    title = category,
    labFormat = labelFormat(prefix = "$"),
    opacity = 1
  ) %>%
  # Layers control
  addLayersControl(
    baseGroups = c("Satellite", "Base"),
    options = layersControlOptions(collapsed = FALSE)
  )
m


```

### Camera activity through time

The `r n.stat` stations have resulted in a total of `r as.character(round(sum(eff$Days, na.rm=T),0))` camera days (mean = `r round(mean(aggregate(Days~location, data=eff,  FUN=sum, na.rm=T)$Days),1)` days per station; min = `r round(min(aggregate(Days~location, data=eff,  FUN=sum, na.rm=T)$Days),1)`; max = `r round(max(aggregate(Days~location, data=eff,  FUN=sum, na.rm=T)$Days),1)`).The daily break down of camera activity is as follows:

```{r activity, echo=F, fig.height=eff.height}

# Adjust layout
par(mar=c(2,6,1,1))
plot(c(min(eff$deployment_begin_date, na.rm=T), max(eff$deployment_end_date, na.rm=T)), c(1,n.stat), las=1, ylab="", xlab="", type="n", yaxt="n")

# Have the first station plot at the top 
plot.order <- rev(unique(eff$location))

axis(2, at= 1:n.stat, labels= plot.order, las=1, cex.axis=0.4)
#mtext("Camera Deployment ID", 2, 4)
# Make lines for each of the cameras
for(i in 1:length(plot.order))
{
  abline(h=i, col=rgb(0,0,0,0.1))
  tmp <- eff[eff$location==plot.order[i],]
  for(j in 1:nrow(tmp))
    {
      lines(c(tmp$deployment_begin_date[j],
                       tmp$deployment_end_date[j]),
            c(i,i), lwd=2)
    }
  
}

```

Figure 2: Where black lines denote a camera which is active, white space indicates cameras which are inactive.

## Raw camera detections

To date, there have been `r nrow(img)` image classifications, `r nrow(dat)` of which are classified as containing identified species (`r round((nrow(dat)/nrow(img))*100,1)`%).

Of the detections which have been identified, there are `r length(levels(factor(dat$common_name)))` species.

```{r captures, echo=F, fig.height=sp.height}
layout(matrix(c(1,1,2), 1, 3, byrow = TRUE))
det.sum.total <- as.data.frame(count(dat, common_name))
det.sum.total <- det.sum.total[order(det.sum.total$n),]

par(mar=c(5,16,1,1))
barplot(det.sum.total$n, names.arg = paste0(det.sum.total$common_name, 
                                           " (n =", det.sum.total$n,")")   , las=1, cex.names=1, xlab="Total detections", horiz=T)
i <-1
for(i in 1:nrow(det.sum.total))
{
  tmp <- subset(dat, common_name==det.sum.total$common_name[i])
  det.sum.total$Locations[i] <- length(unique(tmp$location))
}
par(mar=c(5,1,1,1))

barplot(det.sum.total$Locations/n.stat, las=1, cex.names=0.7, xlab="Proportion of sites detected", horiz=T, xlim=c(0,1))
abline(v=1, lty=2)

```

## Detection check

The following plot helps you determine if you have detections occuring outside of the times cameras are active. *Important note* You can still get detections outside of the activity period if you have decided that the field of view was shifted and the data is un-comparable to that which was collected earlier.

```{r, include=F}
# Make species colour codes
tmp3 <- data.frame("common_name"=unique(dat$common_name),"Colour"= wheel("lightseagreen", num = length(unique(dat$common_name))))

```

```{r detecion summary, echo=F, message=F, warning=F}

# To do this make a plot dataframe
tmp4 <- data.frame("location"=plot.order, "Plot.grp"=ceiling(1:length(unique(eff$location))/20))

eff <- left_join(eff,tmp4, by="location")

j <- 1
i <- 1
for(j in 1:length(unique(eff$Plot.grp)))
{
    layout(matrix(c(1,1,1,2), 1, 4, byrow = TRUE))
    par(mar=c(2,8,1,1))
    
    plot(c(min(eff$deployment_begin_date, na.rm=T), max(eff$deployment_end_date, na.rm=T)),      c(1,length(unique(eff$location[eff$Plot.grp==j]))), las=1, ylab="", xlab="", type="n", yaxt="n")
    
    axis(2, at= 1:length(unique(eff$location[eff$Plot.grp==j])), labels= unique(eff$location[eff$Plot.grp==j]), las=1, cex.axis=1)
    # Make lines for each of the cameras
    for(i in 1:length(unique(eff$location[eff$Plot.grp==j])))
    {
      abline(h=i, col=rgb(0,0,0,0.1))
      tmp <- eff[eff$location==unique(eff$location[eff$Plot.grp==j])[i],]
      
      tmp2 <- dat[dat$location==tmp$location[1],]
      tmp2 <- left_join(tmp2, tmp3)
      points(tmp2$date_detected, rep(i,nrow(tmp2)), pch="|", col= tmp2$Colour)
    
      for(k in 1:nrow(tmp))
        {
          lines(c(tmp$deployment_begin_date[k],
                           tmp$deployment_end_date[k]),
                c(i,i), lwd=2)
        }
      }
    par(mar=c(0,0,1,0))
    plot.new()
    legend("topleft", legend=tmp3$common_name, fill=tmp3$Colour, xpd=TRUE, cex=1.1 )

}

```


Using an independance threshold of `r independent` minutes, the number of detections is reduced to `r nrow(ind.dat)`. The rest of the analyses are conducted with this data. The summary of detections is as follows:

```{r ind captures, echo=F, fig.height=sp.height, eval=T}

layout(matrix(c(1,1,2), 1, 3, byrow = TRUE))
det.sum.total <- as.data.frame(count(ind.dat, common_name))
det.sum.total <- det.sum.total[order(det.sum.total$n),]

par(mar=c(5,16,1,1))
barplot(det.sum.total$n, names.arg = paste0(det.sum.total$common_name,
                                           " (n =", det.sum.total$n,")"), las=1, cex.names=1, xlab="Total detections", horiz=T)
i <-1
for(i in 1:nrow(det.sum.total))
{
  tmp <- subset(ind.dat, common_name==det.sum.total$common_name[i])
  det.sum.total$Locations[i] <- length(unique(tmp$location))
}
par(mar=c(5,1,1,1))

barplot(det.sum.total$Locations/n.stat, las=1, cex.names=0.7, xlab="Proportion of sites detected", horiz=T, xlim=c(0,1))
abline(v=1, lty=2)

```

## Species metadata

Of the images classfied as containing animals, the proportion of photographs assigned to the following catagories are as follows:

### Sex

```{r sex, echo=F, include=F}


b<-ind.dat %>% group_by(common_name, sex) %>% summarise(n = sum(group_count))
d<-t(spread(b, sex, n))
e<-as.data.frame(d[2:4,])
names(e)<-d[1,]
e[is.na(e)]<-0
e
str(e)

tmp <- e
for (i in 1: ncol(e)){
  tmp[,i] <- as.numeric(e[,i])
}



dat[,"sex"] <- factor(dat[,"sex"])
cols <- wheel("lightseagreen", num = length(levels(dat[,"sex"])))

data_percentage <- apply(tmp, 2, function(x){x*100/sum(x,na.rm=T)})

```

```{r sex plot, echo=F, fig.height=sp.height}
layout(matrix(c(1,1,1,2), 1, 4, byrow = TRUE))
par(mar=c(5,10,1,1))
barplot(data_percentage , border="white",col= cols, ylab="", las=1, xlab="% of observations", cex.names=0.7, horiz=2)
par(mar=c(0,0,4,0))
plot.new()
legend("topleft", legend=row.names(tmp), fill=cols, xpd=TRUE, cex=1.1 )
```

### Age

```{r age, echo=F, include=F}

str(ind.dat)

b<-ind.dat %>% group_by(common_name, age_class) %>% summarise(n = sum(group_count))
d<-t(spread(b, age_class, n))
e<-as.data.frame(d[2:4,])
names(e)<-d[1,]
e[is.na(e)]<-0
e
str(e)

tmp <- e
for (i in 1: ncol(e)){
  tmp[,i] <- as.numeric(e[,i])
}


dat[,"age_class"] <- factor(dat[,"age_class"])
cols <- wheel("lightseagreen", num = length(levels(dat[,"age_class"])))


data_percentage <- apply(tmp, 2, function(x){x*100/sum(x,na.rm=T)})

```

```{r age plot, echo=F, fig.height=sp.height}
layout(matrix(c(1,1,1,2), 1, 4, byrow = TRUE))
par(mar=c(5,10,1,1))
barplot(data_percentage , border="white",col= cols, ylab="", las=1, xlab="% of observations", cex.names=0.7, horiz=2)
par(mar=c(0,0,4,0))
plot.new()
legend("topleft", legend=row.names(tmp), fill=cols, xpd=TRUE, cex=1.1 )

```

## Group size distribution

```{r group size, echo=F, eval=T,fig.height=sp.height}

ind.dat2<-ind.dat %>%
  group_by(event_id, common_name) %>%
  summarise(gc = sum(group_count))


par(mfrow=c(1,1))
par(mar=c(5,10,1,1))
plot(jitter(as.numeric(as.factor(ind.dat2$common_name)))~jitter(as.numeric(ind.dat2$gc)), xlab="Minimum group size", yaxt="n", las=1, ylab="")
axis(2, 1:length(unique(ind.dat2$common_name)), labels=levels(as.factor(ind.dat2$common_name)), las=2, cex.axis=0.6)

```

# Site-level species covariance

This plot shows the covariance between different species at the site level for species with \>5 unique detections. For example, if you typically get lots of caribou and bears at the same site, they will have positive covariance. If you get caribou where you don't get bears, they will have negative covariance.

```{r covariance, echo=F, fig.height=sp.height,fig.width=sp.height, eval=T}
par(mfrow=c(1,1))
tmp <- as.data.frame.matrix(table(ind.dat$location, ind.dat$common_name))
tmp <- tmp[colSums(tmp)>5]
M <- cor(tmp)

corrplot(M, method="color", #col=matrix(col(200)),
         type="upper", order="hclust",
         # addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, tl.cex = 0.7, #Text label color and rotation
         # Combine with significance
         # p.mat = p.mat, sig.level = 0.01, insig = "blank",
         # hide correlation coefficient on the principal diagonal
         diag=FALSE
         )

```

## Site-level temporal plots

### Summary

Across all sites and species:

```{r calculate monthly data summaries, echo=F, eval=T, include= F}
# Capture rates through time
focal.sp <- as.character(det.sum.total[det.sum.total$n>0,]$common_name)
focal.sp <- focal.sp[order(focal.sp)]
# Remove any blanks
focal.sp <- focal.sp[focal.sp!=""]

# We need to know how many detections there are in each month -> create a row lookup
# This is just a list of every day a camera was active.

tmp <- eff[is.na(eff$deployment_end_date)==F,]
daily.lookup <- list()
for(i in 1:nrow(tmp))
{
  if(as.Date(tmp$deployment_begin_date[i])!=as.Date(tmp$deployment_end_date[i]))
  {
    daily.lookup[[i]] <- data.frame("Date"=seq(as.Date(tmp$deployment_begin_date[i])+1, as.Date(tmp$deployment_end_date[i]), by="days"), "location"=tmp$location[i])
  }
}
row.lookup <- do.call(rbind, daily.lookup)

# Now determine capture rates
# Make a data frame by month and year
mon.dat <- unique(substr(ind.dat$date_detected, 1,7))
mon.dat <- data.frame("Month"=mon.dat[order(mon.dat)], "Effort"= NA)
mon.dat[as.character(focal.sp)] <- NA
i<-1
for(i in 1:nrow(mon.dat))
{
  mon.dat$Effort[i] <- nrow(subset(row.lookup, substr(row.lookup$Date,1,7)==mon.dat$Month[i]))
  mon.dat$Total.CR[i] <- (nrow(subset(ind.dat, substr(ind.dat$date_detected,1,7)==mon.dat$Month[i]))/mon.dat$Effort[i])*100
}

for(i in 1:length(focal.sp))
{
  for(j in 1:nrow(mon.dat))
  {
    tmp <- subset(ind.dat, common_name==as.character(focal.sp)[i] & substr(ind.dat$date_detected,1,7)==mon.dat$Month[j])
    mon.dat[j, as.character(focal.sp[i])] <- (nrow(tmp)/mon.dat$Effort[j])*100
  }
}

mon.dat$timestamp <- strptime(paste0(as.character(mon.dat$Month),"-15"), "%Y-%m-%d")

# Remove any silly values 
mon.dat <- mon.dat[is.infinite(mon.dat$Total.CR)==F,]

```

```{r overall CR, echo=F, fig.height=4, eval=T}

par(mfrow=c(1,2))

xtick <- as.POSIXct(paste0(year(mon.dat$timestamp), "-", str_pad(month(mon.dat$timestamp), 2, side = "left", pad = "0"), "-01" ))



plot(mon.dat$timestamp, mon.dat$Effort, ylab="Monthly Effort (days)", xlab="Date", type="l", las=2)
points(mon.dat$timestamp, mon.dat$Effort, pch=19, col=rgb(0,0,0,0.4))
axis(side=1, at=xtick, labels = FALSE )

# Overall capture rate
plot(mon.dat$timestamp, mon.dat$Total.CR, ylab="Monthly total CR per 100 days", xlab="Date", type="l", las=2, ylim=c(0, max(mon.dat$Total.CR)))
points(mon.dat$timestamp, mon.dat$Total.CR, pch=19, col=rgb(0,0,0,0.4))
axis(side=1, at=xtick, labels = FALSE)

```

### Species-specific temporal trends

Species level variation in monthly capture rates are as follows:

```{r, echo=F, eval=T}
par(mfrow=c(2,3))
for(i in 1:length(focal.sp))
{
  plot(mon.dat$timestamp, mon.dat[,as.character(focal.sp)[i]], ylab="Capture Rate per 100 days", xlab="", type="l", las=2, main=focal.sp[i])
  points(mon.dat$timestamp, mon.dat[,as.character(focal.sp)[i]], pch=19, col=rgb(0,0,0,0.4))
  axis(side=1, at=xtick, labels = FALSE )
}

```

## Species-specific maps

Gray circles are stations where there are no detections of that species. Blue circles are stations where that species was detected. The size of the circle increases with the number of independent detections, including group size.

```{r, eval=T, echo=F, message=FALSE, warning=FALSE}

# get rid of NA counts because they F up the script here
spatdata<-spatdata[!is.na(spatdata$group_count),]
species<-unique(spatdata$common_name)

i<-1

for (i in 1:length(species)){
  tmp<-spatdata[spatdata$common_name==species[i],]
  tmp2<-tmp%>%group_by(common_name, location) %>% summarise(n= sum(group_count))
  tmp2$n<-as.numeric(tmp2$n)
  
  x=tm_shape(tuyeta) +
    tm_borders() +
    tm_layout(
      title = species[i],
      title.position = c("center", "top"),
      inner.margins = c(0.07, 0.03, 0.15, 0.03)) +
    tm_shape(spatdata) +
    tm_symbols(size = 0.2, col="lightgrey") +
    tm_shape(tmp2) +
    tm_bubbles("n",
               col ='blue',
               breaks = c(1, 5, 10, max(tmp$n)),
               size.max = 10,
               sizes.legend= c(1, 5, 10, if(max(tmp2$n)>10) max(tmp2$n))) +
    tm_layout(legend.position = c("right","bottom"))
    
print(x)

}

```
