```{r setup and tests, include=FALSE}

## README FIRST ##
#Read and run this chunk of code line by line in the R Console (do not press knit) - there are some questions below which you will have to answer and some logic tests to complete. Once you are happy that the conditions have been satisfied, hit 'knit' above. 

# Load required packages

library(leaflet)
library(dplyr)
library(colortools)
library(corrplot)
library(lubridate)
library(sf)
library(tmap)
library(stringr)
library(tidyr)
library(fishmethods)

# Load your data [change the files paths to your data locations]
# eff - recordings transcribed (effort)
# sta - station covariates
# dat - detection data 
# ind - independent detection data
# spp - species list
# tuyeta - shapefile outline of the study area
# spatdata - independent detections joined spatially with lat longs

project<-"Tuyeta"

# Timezone [Use UTC if your ARUs do not correct for daylight saving time, if they do use the timezone where the data was collected]
tz <- "UTC"

category <- "project" # for summary graphs
colour <- "lightseagreen" # Define a colour from the R options to base the colourscheme

eff <- read.csv("../1. Master Data/Tuyeta_ARU_Deployment_Data_v5.csv")

sta <- read.csv("../1. Master Data/Tuyeta_Station_Covariates_v5.csv")
sta<-droplevels(sta[sta$location %in% eff$location,]) #keep only the stations with ARU data

dat <- read.csv("../1. Master Data/Tuyeta_ARU_Detection_Data_v5.csv")

spp <- read.csv("../1. Master Data/Tuyeta_ARU_Species_List_v5.csv")

ind <- read.csv("../1. Master Data/Tuyeta_ARU_Independent_Detections_v5.csv")
str(ind)

tuyeta <- st_read("../6. Spatial data/Tuyeta/Tuyeta_boundary/Tuyeta.shp")
tuyeta <- st_transform(tuyeta, crs = 3580)

spatdata=st_read("../1. Master data/Tuyeta_Station_Covariates_v5.csv",
        options=c("X_POSSIBLE_NAMES=longitude","Y_POSSIBLE_NAMES=latitude"))

spatdata<- merge(spatdata, ind, by = "location")

```

```{r non-adjustable options, echo=F, include=F}


# Prepare dates
eff$recording_begin <- as.POSIXct(strptime(eff$recording_begin, "%Y-%m-%d %H:%M", tz ))

# Count the number of ARU stations
n.stat <- length(unique(eff$location))
n.stat

# Generate colours to display the category levels - R needs them as a factor
sta$project <- as.factor(sta$project)
col.cat <- wheel(colour, num = length(levels(sta[,"project"])))
sta$Cols <- col.cat[sta[,"project"]]

# How big should the figures be
eff.height <- 8
if(length(unique(eff$location))>80)
   {
     eff.height <- length(unique(eff$location))/10
   }

sp.height <- 7
if(length(unique(ind$common_name))>20)
   {
     sp.height <- 7+(length(unique(ind$common_name))/8)
   }


```


# `r project` Project

### ARU locations

To date there have been ARU deployments at `r n.stat` unique locations.

```{r map, echo=F}

m <- leaflet() %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data
  addProviderTiles(providers$Esri.WorldTopoMap, group="Base") %>%     
  addCircleMarkers(lng=sta$longitude, lat=sta$latitude,
                   color=sta$Cols,
                   popup=paste(sta$location, sta[,"project"])) %>%
 addLegend("bottomleft", colors = col.cat,  labels = levels(sta[,"project"]),
    title = category,
    labFormat = labelFormat(prefix = "$"),
    opacity = 1
  ) %>%
  # Layers control
  addLayersControl(
    baseGroups = c("Satellite", "Base"),
    options = layersControlOptions(collapsed = FALSE)
  )
m


```


### ARU activity through time

The `r n.stat` stations have resulted in a total of  `r nrow(eff)` recordings (mean = `r round(nrow(eff)/n.stat)` recordings per station; min = `r eff %>% group_by(location) %>% summarise(n=n()) %>% summarise(min(n)) %>% unname()`; max = `r eff %>% group_by(location) %>% summarise(n=n()) %>% summarise(max(n)) %>% unname()`).The daily break down of ARU activity is as follows:

```{r activity, echo=F, fig.height=eff.height}

# Adjust layout
par(mar=c(2,6,1,1))
plot(c(min(eff$recording_begin, na.rm=T), max(eff$recording_begin, na.rm=T)),
     c(1,n.stat), las=1, ylab="", xlab="", type="n", yaxt="n")

# Have the first station plot at the top 
plot.order <- sort(unique(eff$location),decreasing = T)

axis(2, at= 1:n.stat, labels= plot.order, las=1, cex.axis=0.4)

for(i in 1:length(plot.order))
{
  abline(h=i, col=rgb(0,0,0,0.1))
  tmp <- eff[eff$location==plot.order[i],]
  for(j in 1:nrow(tmp))
  {
    points(tmp$recording_begin[j],
          i, lwd=2, pch = 19, cex = 0.8)
  }
}

```

## ARU activity by time of day

```{r time of day, echo = F, fig.height = 5, warning = F, message = F}
par(mar=c(5.1, 4.1, 4.1, 2.1))
eff$hour = hour(eff$recording_begin)
hist(eff$hour, xlim = c(0,24), breaks = 24, main = "Number of recordings at each hour of the day (0 - 24)", xlab = "Hour")

eff <- merge(eff,sta[,1:3], by = "location")

day <- day(eff$recording_begin)
month <- month(eff$recording_begin)
year <- year(eff$recording_begin)
hour <- hour(eff$recording_begin)
timezone <- rep(-6, nrow(eff))
lat<- eff$latitude
lon = eff$longitude

sundata = astrocalc4r(day = day, month = month, year = year, hour = hour, timezone = timezone, lat = lat, lon = lon)
sundata = cbind(eff, sundata)
sundata$sun_angle = 90 - sundata$zenith

sundata <- sundata %>% mutate(type = case_when(sun_angle < -18 ~ "night",
    sun_angle >= -18 & sun_angle < -12 ~ "astronomical twilight",
    sun_angle >= -12 & sun_angle < -6 ~ "nautical twilight",
    sun_angle >= -6 & sun_angle < 0 ~ "civil twilight",
    sun_angle >=0 ~ "day"
                   ))
```

```{r get breaks, eval = T, include = F}

hist_breaks <- hist(sundata$sun_angle)$breaks
```

```{r sun angle histrogram, echo = F, message = F, warning = F, fig.height = 5}

#define colors to use in histogram based on break points
color_list <- rep("#D1E5F0", length(hist_breaks)) # day
color_list[hist_breaks < 0] <- "#4393C3" # twilight
color_list[hist_breaks < -18] <- "#053061" # night

hist(sundata$sun_angle, main = "Number of recordings at different sun angles", xlab = "Sun altitude, degrees", ylab = "Number of recordings", col = color_list)
legend("topright",c("Day","Twilight","Night"), fill = c("#D1E5F0","#4393C3","#053061" ))

```

## Raw ARU detections

To date, there have been `r nrow(dat)` sound classifications.

Of the detections which have been identified, there are `r nrow(spp)` species.

```{r captures, echo=F, fig.height=sp.height}
layout(matrix(c(1,1,2), 1, 3, byrow = TRUE))
det.sum.total <- as.data.frame(count(dat, common_name))
det.sum.total <- det.sum.total[order(det.sum.total$n),]
det.sum.total <- det.sum.total %>% filter(common_name %in% spp$common_name)

par(mar=c(5,16,1,1))
barplot(det.sum.total$n, names.arg = paste0(det.sum.total$common_name, 
                                           " (n =", det.sum.total$n,")")   , las=1, cex.names=1, xlab="Total detections", horiz=T)
i <-1
for(i in 1:nrow(det.sum.total))
{
  tmp <- subset(dat, common_name==det.sum.total$common_name[i])
  det.sum.total$Locations[i] <- length(unique(tmp$location))
}
par(mar=c(5,1,1,1))

barplot(det.sum.total$Locations/n.stat, las=1, cex.names=0.7, xlab="Proportion of sites detected", horiz=T, xlim=c(0,1))
abline(v=1, lty=2)

```


## Independent Detections

Assuming closure during the breeding season (i.e., subsequent detections of a species within the same year at a given location is assumed to be the same individual), the number of detections is reduced to `r nrow(ind)`. The rest of the analyses are conducted with this data. The summary of detections is as follows:

```{r ind captures, echo=F, fig.height=sp.height, eval=T}

layout(matrix(c(1,1,2), 1, 3, byrow = TRUE))
det.sum.total <- as.data.frame(count(ind, common_name))
det.sum.total <- det.sum.total[order(det.sum.total$n),]
det.sum.total <- det.sum.total %>% filter(common_name %in% spp$common_name)


par(mar=c(5,16,1,1))
barplot(det.sum.total$n, names.arg = paste0(det.sum.total$common_name,
                                           " (n =", det.sum.total$n,")"), las=1, cex.names=1, xlab="Total detections", horiz=T)
i <-1
for(i in 1:nrow(det.sum.total))
{
  tmp <- subset(ind, common_name==det.sum.total$common_name[i])
  det.sum.total$Locations[i] <- length(unique(tmp$location))
}
par(mar=c(5,1,1,1))

barplot(det.sum.total$Locations/n.stat, las=1, cex.names=0.7, xlab="Proportion of sites detected", horiz=T, xlim=c(0,1))
abline(v=1, lty=2)

```


## Species-specific maps

Gray circles are stations where there are no detections of that species. Blue circles are stations where that species was detected. The size of the circle increases with the number of independent detections, including group size.

```{r eval=T, echo=F, message=FALSE, warning=FALSE}

# get rid of NA counts because they F up the script here
spatdata<-spatdata[!is.na(spatdata$abundance),]
species<-spp$common_name

i<-1
for (i in 1:length(species)){
  tmp<-spatdata[spatdata$common_name==species[i],]
  tmp2<-tmp%>%group_by(common_name, location) %>% summarise(n= sum(abundance))
  tmp2$n<-as.numeric(tmp2$n)
  
  x=tm_shape(tuyeta) +
    tm_borders() +
    tm_layout(
      title = species[i],
      title.position = c("center", "top"),
      inner.margins = c(0.07, 0.03, 0.15, 0.03)) +
    tm_shape(spatdata) +
    tm_symbols(size = 0.2, col="lightgrey") +
    tm_shape(tmp2) +
    tm_bubbles("n",
               col ='blue',
               breaks = c(1, 5, 10, max(tmp$n)),
               size.max = 10,
               sizes.legend= c(1, 5, 10, if(max(tmp2$n)>10) max(tmp2$n))) +
    tm_layout(legend.title.size= 0.1, legend.position = c("right","bottom"))
    
print(x)

}

```
